---
title: "CV on real data"
author: "Halvard"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Copy relevant code from the other files
First, lets copy the necessary code from the file "WinBUGS and the adaptive multivariate model.Rmd", as importing .Rmd files is problematic, the important parts of the code will be copied in here:
```{r}
library(readr)
library(ggplot2)

library(sf)
library(spdep)
library(dplyr)

ExpectedCases <- readRDS("ExpectedCasesMiguel.rds")
ObservedCases <- readRDS("ObservedCases.rds")
PopulationData <- readRDS("PopulationData.rds")
nDiseasesTot <- as.numeric(ncol(ObservedCases)) #the number of causes

shp_prov <- st_read("Data/Provinces/Provincias_ETRS89_30N.shp")
colnames(shp_prov) <- c("Index", "ProvName", "ProvName1", "CCAA_Index", "CCAA_Name", "geometry")
CCAA_to_remove <- c("Illes Balears", "Canarias", "Ceuta", "Melilla")
shp_prov <- filter(shp_prov, !CCAA_Name %in% CCAA_to_remove)
carto <- shp_prov

#Spatial structure
carto.nb <- poly2nb(carto) # Neighbours list of each geographic unit with class nb
carto.wb <- nb2WB(carto.nb) # List with the adjacency vector (carto.wb$adj) and
index <- c(1, cumsum(carto.wb$num)) # Vector to identify the positions of the neighbors

ss_provinces = read_sf("Data/Provinces/Provincias_ETRS89_30N.shp")

ss_ar = read_sf("Data/AutonomousRegions/Comunidades_Autonomas_ETRS89_30N.shp")

# remove provinces outside mainland Spain

out = c("Islas Baleares", "Las Palmas", "Santa Cruz de Tenerife", "Ceuta", "Melilla")

out2 = c("Canarias", "Islas Baleares", "Ceuta", "Melilla")

ss_provinces = ss_provinces |> filter(!(Texto %in% out))

ss_ar = ss_ar %>% filter(!(Texto %in% out2))

nAreas <- length(carto.wb$num)
nEdges <- length(carto.wb$adj)/2
```



Secondly, import the necessary code for the Border weighted ICAR, again copied in:
```{r}
library(INLA)

inla.rgeneric.BWICAR = function(
  cmd = c("graph", "Q", "mu", "initial", "log.norm.const","log.prior", "quit"),
  theta = NULL)
{
#Input:
#df: contains the relevant information for the areal data
#prior_str is either Gamma, PC1, PC3 or U
#AMat is the adjacency matrix, so 1 for bordering regions representing a first order structure
  
  envir = parent.env(environment())
  
  interpret_theta <- function() { return(list(tau1 = exp(theta[1L]), 
                                              tau2 = exp(theta[2L])))}
  
  graph <- function() {return(Q())}
  
  Q <- function() {
    N <- nrow(df) #df is passed as an argument, N is the number of regions
    R1 <- matrix(0, nrow = N, ncol = N) #non-conflict
    R2 <- matrix(0, nrow = N, ncol = N) #conflict
    
    #assume the adjacency matrix is called AMat, then
    non_zero_indices <- which(AMat == 1, arr.ind = TRUE)
    
    for (k in seq_len(nrow(non_zero_indices))) { #nrow?
      i <- non_zero_indices[k, 1]
      j <- non_zero_indices[k, 2]
      
      if(df[["Cod_CCAA"]][i] == df[["Cod_CCAA"]][j]){
        R1[i, j] <- -1
        R1[i, i] <- R1[i, i] + 1
      }
      else{
        R2[i, j] <- -1
        R2[i, i] <- R2[i, i] + 1
      }
    }
    
    gv <- exp(1 / N * sum(log(diag(INLA:::inla.ginv(R1 + R2))))) #scaling constant
    R_star_list <- list(R1 = R1*gv, R2 = R2*gv)
    
    p <- interpret_theta()
    Q <- R_star_list$R1 * p$tau1 + R_star_list$R2 * p$tau2
    return(inla.as.sparse(Q)) #sparse representation
  }
    
  mu <- function() {return(numeric(0))}
  
  initial <- function() {return(c(4, 4))}#Default initial for precisions is 4
  
  log.norm.const <- function() {return(numeric(0))}
  
  # log.prior <- function() {
  #   p <- interpret_theta()
  #   prior <- inla.pc.dprec(p$tau1, u = 1, alpha = 0.01, log=TRUE) + log(p$tau1) +
  #          inla.pc.dprec(p$tau2, u = 1, alpha = 0.01, log = TRUE) + log(p$tau2)
  #   return(prior)
  # }
  
  log.prior <- function() {#default: shape = 1, rate = 0.00005 for tau
    p <- interpret_theta()
    if (pr_str == "gamma")  {
      prior <- dgamma(p$tau1, shape = 1, rate = 0.00005, log = TRUE) + log(p$tau1) +
        dgamma(p$tau2, shape = 1, rate = 0.00005, log = TRUE) + log(p$tau2)}
    else if (pr_str == "PC3") {
      prior <- inla.pc.dprec(p$tau1, u = 3, alpha = 0.05, log=TRUE) + log(p$tau1) +
        inla.pc.dprec(p$tau2, u = 3, alpha = 0.05, log=TRUE) + log(p$tau2)}
    else if (pr_str == "PC1") {
      prior <- inla.pc.dprec(p$tau1, u = 1, alpha = 0.01, log=TRUE) + log(p$tau1) +
        inla.pc.dprec(p$tau2, u = 1, alpha = 0.01, log=TRUE) + log(p$tau2)}
    else if (pr_str == "U") {prior <- -0.5*log(p$tau1) - 0.5*log(p$tau2)}
    return(prior)
  }
  
  quit <- function() {return(invisible())}
  
  #to ensure theta is defined
  if (!length(theta)) theta = initial()
  
  vals <- do.call(match.arg(cmd), args = list())
  return(vals)
}

nb <- poly2nb(ss_provinces)
adj_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)

BW_model <- inla.rgeneric.define(inla.rgeneric.BWICAR, df = ss_provinces,
                                    AMat = adj_matrix, pr_str = "PC3")
```


# Implementing the adaptive ICARS with rgeneric
Both EW-ICAR and RW-ICAR can easily compute a list of all the taus for the edges. From this it is possible to construct the precision matrix Q, maybe along with the index set as well. I will start by experimenting with the creation of Q before implementing the whole matrix in rgeneric. After the models are defined in rgeneric I will perform cross validation for all the diseases, and further investigate the scarce and prominent cause of death, maybe also total mortality. The comparison of the models will be done for all n, which is always one for Wakefield, and a choice concerning test diseases for lower n, i.e. 10, 20, 50 must be made in regards to using a known disease or not in the training data.

```{r}
index2 <- index
index2[1] <- 0

#as a function
constructQ <- function(tau, indexes, nAreas, adj){
  #tau is a list of each edge, 222 elements, indexes keeps track of the indexes for the adj list, which contains all the indexes of neighbors. nAreas is the number of areas.
  Q <- matrix(0, nrow = nAreas, ncol = nAreas)
  count_num <- 1
  for (i in 1:nAreas){
    for (j in adj[(indexes[i]+1):indexes[i+1]]){
      Q[i, j] <- -tau[count_num]
      
      Q[i, i] <- Q[i, i] + tau[count_num]
      count_num <- count_num + 1
    }
  }
  return(Q)
}

```

## Making the matrices 
Make the matrices for EW-ICAR and RW-ICAR for different number of diseases n.

```{r}
edges_df <- readRDS("Edges_df//Edges_W_scaling_WO_theta")

Q_RW10 <- constructQ(edges_df$edge_value_RW_10, index2, nAreas, carto.wb$adj)
Q_RW20 <- constructQ(edges_df$edge_value_RW_20, index2, nAreas, carto.wb$adj)
Q_RW50 <- constructQ(edges_df$edge_value_RW_50, index2, nAreas, carto.wb$adj)
Q_RW86 <- constructQ(edges_df$edge_value_RW_86, index2, nAreas, carto.wb$adj)
Q_RW_Cancer <- constructQ(edges_df$edge_value_RW_cancer, index2, nAreas, carto.wb$adj)

Q_EW10 <- constructQ(edges_df$edge_value_EW_10, index2, nAreas, carto.wb$adj)
Q_EW20 <- constructQ(edges_df$edge_value_EW_20, index2, nAreas, carto.wb$adj)
Q_EW50 <- constructQ(edges_df$edge_value_EW_50, index2, nAreas, carto.wb$adj)
Q_EW86 <- constructQ(edges_df$edge_value_EW_86, index2, nAreas, carto.wb$adj)
Q_EW_Cancer <- constructQ(edges_df$edge_value_EW_cancer, index2, nAreas, carto.wb$adj)

```

## The adaptive ICAR as a rgeneric definiton
The function defining the latent term with a given structure matrix $Q$ as an input. I think this could be the same for each method, simply different structure matrices provided. If they differ, it should only be for priors??

Should I scale with inla.model.scale = T, or scale it manually inside the rgenric definition in the Q() function?? -- I think it must be manual, before scaling with tau.

Is it fine with numeric(0) for the log constant??

```{r}
inla.rgeneric.Validation = function(
  cmd = c("graph", "Q", "mu", "initial", "log.norm.const","log.prior", "quit"),
  theta = NULL)
{
  #Input:
  #Q as the unscaled precision structure matrix
  #nAreas is the number of provinces
  #pr_str indicates which prior to use
  
  envir = parent.env(environment())
  
  interpret_theta <- function() {return(list(tau = exp(theta[1L])))}
  
  graph <- function() {return(Q())}
  
  Q <- function() {
    p <- interpret_theta()
    
    gv <- exp(1 / nAreas * sum(log(diag(INLA:::inla.ginv(R))))) #the geometric variance
    return(inla.as.sparse(p$tau * gv * R))
  }
  
  mu <- function() {return(numeric(0))}
  
  initial <- function() {return(4)}#default for precisions: initial = 4
  
  log.norm.const <- function() {return(numeric(0))} #Inla computes it, 
  
  log.prior <- function() {#default: shape = 1, rate = 0.00005 for tau
    p <- interpret_theta()
    if (pr_str == "gamma")  {prior <- dgamma(p$tau, shape = 1, rate = 0.00005, log = TRUE) + log(p$tau)}
    else if (pr_str == "PC") {prior <- inla.pc.dprec(p$tau, u = 3, alpha = 0.05, log=TRUE) + log(p$tau)}
    else if (pr_str == "U") {prior <- -0.5*log(p$tau)}
    return(prior)
  }
  
  quit <- function() {return(invisible())}
  
  #to ensure theta is defined
  if (!length(theta)) theta = initial()
  
  vals <- do.call(match.arg(cmd), args = list())
  return(vals)
}

#The specific models

#prior_str <- "PC"
#prior_str <- "gamma"
prior_str <- "U"

RW10_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_RW10, nAreas = nAreas, pr_str = prior_str)
RW20_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_RW20, nAreas = nAreas, pr_str = prior_str)
RW50_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_RW50, nAreas = nAreas, pr_str = prior_str)
RW86_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_RW86, nAreas = nAreas, pr_str = prior_str)
RW_C_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_RW_Cancer, nAreas = nAreas, pr_str = prior_str)

EW10_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_EW10, nAreas = nAreas, pr_str = prior_str)
EW20_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_EW20, nAreas = nAreas, pr_str = prior_str)
EW50_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_EW50, nAreas = nAreas, pr_str = prior_str)
EW86_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_EW86, nAreas = nAreas, pr_str = prior_str)
EW_C_model <- inla.rgeneric.define(inla.rgeneric.Validation, R = Q_EW_Cancer, nAreas = nAreas, pr_str = prior_str)
```

Then also create the structure matrices R1 and R2 for the BW-ICAR from Wakefield, which will be used in the validation step to construct the scaled precision matrix Q in the validation loop.

```{r}
R1 <- matrix(0, nrow = nAreas, ncol = nAreas) #non-conflict
R2 <- matrix(0, nrow = nAreas, ncol = nAreas) #conflict

#assume the adjacency matrix is called AMat, then
non_zero_indices <- which(adj_matrix == 1, arr.ind = TRUE)

for (k in seq_len(nrow(non_zero_indices))) {
  i <- non_zero_indices[k, 1]
  j <- non_zero_indices[k, 2]

  if(ss_provinces[["Cod_CCAA"]][i] == ss_provinces[["Cod_CCAA"]][j]){
    R1[i, j] <- -1
    R1[i, i] <- R1[i, i] + 1
  }
  else{
    R2[i, j] <- -1
    R2[i, i] <- R2[i, i] + 1
  }
}

#image(R1+R2)

```


## The Cross-Validation
The idea here is to use the Q-matrices obtained from multivariate analysis for the two WinBUGS models, and univariately for the Wakefield model, not actually quite sure how the Wakefield model will work. They will be used as adaptive ICAR models an fitted to the univariate areal data with INLA, the model will look something like this:
$$
Y_{i} = Poisson(\lambda_i) \\
\log(\lambda_i) = \log(E_i) + \mu + \phi_i \\
\boldsymbol{\phi} \sim IGMRF(\mathbf{0}, \mathbf{Q}) 
$$
with the Q matrix assigned from the fitted models in the file "WinBUGS and the adaptive multivariate models.rmd", and they are constructed from the lists further up in this file. Should maybe also add a sum to zero constraint on the $\phi$, currently only a soft sum to zero constraint from mean = numeric(0) in the rgenric definition.

### The testing
Note that `res$cpo$logcpo` gets the log sum of the individual cpo-values, so taking minus and dividing by nAreas gives the same as `-mean(log(res$cpo$cpo))`.
```{r, eval=FALSE}

ExpectedCases <- readRDS("ExpectedCasesMiguel.rds")
ObservedCases <- readRDS("ObservedCases.rds")

#For all diseases
test_indecies <- 1:86 #the diseases to be validated on

criteria_df <- data.frame(matrix(NA, nrow = length(test_indecies), ncol = ((2*4 + 3 + 2)*3)))
colnames(criteria_df) <- c("RW_10_WAIC", "RW_20_WAIC", "RW_50_WAIC", "RW_86_WAIC",
                           "EW_10_WAIC", "EW_20_WAIC", "EW_50_WAIC", "EW_86_WAIC",
                           "RW_10_DIC", "RW_20_DIC", "RW_50_DIC", "RW_86_DIC",
                           "EW_10_DIC", "EW_20_DIC", "EW_50_DIC", "EW_86_DIC",
                           "RW_10_CPO", "RW_20_CPO", "RW_50_CPO", "RW_86_CPO",
                           "EW_10_CPO", "EW_20_CPO", "EW_50_CPO", "EW_86_CPO",
                           "BW_WAIC", "BW_DIC", "BW_CPO", "BW_2_WAIC", "BW_2_DIC", "BW_2_CPO",
                           "ICAR_WAIC", "ICAR_DIC", "ICAR_CPO", 
                           "RW_C_WAIC", "RW_C_DIC", "RW_C_CPO", 
                           "EW_C_WAIC", "EW_C_DIC", "EW_C_CPO")

#Need some df to store results, ie. WAIC, LS, CPO and such, like above!

t0 <- Sys.time()

for (j in test_indecies){
  Y <- as.vector(ObservedCases[, j])
  E <- as.vector(ExpectedCases[, j])
  
  data_df <- data.frame(Y = Y, E = E, ID = 1:nAreas)
  
  # The RW-ICAR by Miguel
  
  formula_RW_10 <- Y ~ offset(log(E)) + f(ID, model = RW10_model) #intercept is automatically included
  formula_RW_20 <- Y ~ offset(log(E)) + f(ID, model = RW20_model)
  formula_RW_50 <- Y ~ offset(log(E)) + f(ID, model = RW50_model)
  formula_RW_86 <- Y ~ offset(log(E)) + f(ID, model = RW86_model)
  formula_RW_C <- Y ~ offset(log(E)) + f(ID, model = RW_C_model)
  
  res_RW_10 <- inla(formula_RW_10, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_RW_20 <- inla(formula_RW_20, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_RW_50 <- inla(formula_RW_50, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_RW_86 <- inla(formula_RW_86, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_RW_C <- inla(formula_RW_C, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  
  # The EW-ICAR by Riddervold/Jo
  
  formula_EW_10 <- Y ~ offset(log(E)) + f(ID, model = EW10_model)
  formula_EW_20 <- Y ~ offset(log(E)) + f(ID, model = EW20_model)
  formula_EW_50 <- Y ~ offset(log(E)) + f(ID, model = EW50_model)
  formula_EW_86 <- Y ~ offset(log(E)) + f(ID, model = EW86_model)
  formula_EW_C <- Y ~ offset(log(E)) + f(ID, model = EW_C_model)
  
  res_EW_10 <- inla(formula_EW_10, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_EW_20 <- inla(formula_EW_20, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_EW_50 <- inla(formula_EW_50, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_EW_86 <- inla(formula_EW_86, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  res_EW_C <- inla(formula_EW_C, family = "poisson", data = data_df,
                    control.compute = list(cpo = T, waic = T, dic = T))
  
  #BW-ICAR by Wakefield
  formula_BW <- Y ~ offset(log(E)) + f(ID, model = BW_model)
  res_BW <- inla(formula_BW, family = "poisson", data = data_df,
                 control.compute = list(cpo = T, waic = T, dic = T))
  #construct BW Q from tau1 and tau2
  #taus <- c(res_BW$summary.hyperpar[, "mean"])
  prec1_marginal <- inla.tmarginal(function(x) exp(x),
                                   res_BW$internal.marginals.hyperpar[[1]])
  prec2_marginal <- inla.tmarginal(function(x) exp(x), 
                                   res_BW$internal.marginals.hyperpar[[2]])
  
  # Compute summary including the mode
  summary_prec1 <- inla.zmarginal(prec1_marginal, silent = TRUE)
  summary_prec2 <- inla.zmarginal(prec2_marginal, silent = TRUE)
  
  # Access mode
  mode_prec1 <- summary_prec1$mean
  mode_prec2 <- summary_prec2$mean
  
  Q_mat <- mode_prec1*R1 + mode_prec2*R2
  BW_2_model <- inla.rgeneric.define(inla.rgeneric.Validation, 
                                     R = Q_mat, nAreas = nAreas, pr_str = prior_str)
  
  formula_BW_2 <- Y ~ offset(log(E)) + f(ID, model = BW_2_model)
  
  res_BW_2 <- inla(formula_BW_2, family = "poisson", data = data_df,
                 control.compute = list(cpo = T, waic = T, dic = T))
  
  #The standard ICAR
  
  if (prior_str == "PC"){
    formula_ICAR <- Y ~ offset(log(E)) + 
      f(ID, model = "besag", graph = adj_matrix, scale.model = TRUE,
      hyper = list(prec = list(prior = "pc.prec", param = c(3, 0.05),log = TRUE)))
  }
  else if (prior_str == "gamma"){
    formula_ICAR <- Y ~ offset(log(E)) + 
      f(ID, model = "besag", graph = adj_matrix, scale.model = TRUE)
  }
  else if (prior_str == "U"){
    sdunif="expression: logdens=-log_precision/2; return(logdens)"
    formula_ICAR <- Y ~ offset(log(E)) + 
      f(ID, model = "besag", graph = adj_matrix, scale.model = TRUE, 
        hyper = list(prec = list(prior = sdunif)))
  }
  
  res_ICAR <- inla(formula_ICAR, family = "poisson", data = data_df,
                   control.compute = list(cpo = T, waic = T, dic = T))
  
  #Writing the results to a dataframe
  vec_waic_dic <- c(res_RW_10$waic$waic, res_RW_20$waic$waic, res_RW_50$waic$waic,
                    res_RW_86$waic$waic, res_EW_10$waic$waic, res_EW_20$waic$waic, 
                    res_EW_50$waic$waic, res_EW_86$waic$waic, res_RW_10$dic$dic, 
                    res_RW_20$dic$dic, res_RW_50$dic$dic, res_RW_86$dic$dic, 
                    res_EW_10$dic$dic, res_EW_20$dic$dic, res_EW_50$dic$dic,
                    res_EW_86$dic$dic)
  vec_cpo <- c(-mean(log(res_RW_10$cpo$cpo)), -mean(log(res_RW_20$cpo$cpo)),
               -mean(log(res_RW_50$cpo$cpo)), -mean(log(res_RW_86$cpo$cpo)),
               -mean(log(res_EW_10$cpo$cpo)), -mean(log(res_EW_20$cpo$cpo)),
               -mean(log(res_EW_50$cpo$cpo)), -mean(log(res_EW_86$cpo$cpo)))
  #-mean(log(res_RW1$cpo$cpo))
  vec_BW <- c(res_BW$waic$waic, res_BW$dic$dic, -mean(log(res_BW$cpo$cpo)),
              res_BW_2$waic$waic, res_BW_2$dic$dic, -mean(log(res_BW_2$cpo$cpo)))
  vec_ICAR <- c(res_ICAR$waic$waic, res_ICAR$dic$dic, -mean(log(res_ICAR$cpo$cpo)))
  
  vec_cancer <- c(res_RW_C$waic$waic, res_RW_C$dic$dic, -mean(log(res_RW_C$cpo$cpo)),
                  res_EW_C$waic$waic, res_EW_C$dic$dic, -mean(log(res_EW_C$cpo$cpo)))
  
  criteria_df[j,] <- c(vec_waic_dic, vec_cpo, vec_BW, vec_ICAR, vec_cancer)
}

#runtimes and saving
CV_runtime <- Sys.time() - t0
CV_runtime

#testing
Y <- as.vector(ObservedCases[, 60])
E <- as.vector(ExpectedCases[, 60])
 
data_df <- data.frame(Y = Y, E = E, ID = 1:nAreas)

```


```{r}
# save the criteria_df
prior_str

#saveRDS(criteria_df, file = "CV-Results//CV_no_iidTheta_gamma_W_C_final.rds")

#saveRDS(criteria_df, file = "CV-Results//CV_no_iidTheta_PC_W_C_final.rds")

#saveRDS(criteria_df, file = "CV-Results//CV_no_iidTheta_U_W_C_final.rds")
```


# Make the box-plots for the "CV" results

Plot for all the diseases without a reference level.
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggpubr)

#res_df <- readRDS(file = "CV-Results//CV_no_iidTheta_gamma_W_C_final.rds")

#res_df <- readRDS(file = "CV-Results//CV_no_iidTheta_PC_W_C_final.rds")

res_df <- readRDS(file = "CV-Results//CV_no_iidTheta_U_W_C_final.rds")

index_WAIC <- c(1, 2, 3, 4, 5, 6, 7, 8, 25, 28, 31, 34, 37)

index_DIC <- c(9, 10, 11, 12, 13, 14, 15, 16, 26, 29, 32, 35, 38)

index_CPO <- c(17, 18, 19, 20, 21, 22, 23, 24, 27, 30, 33, 36, 39)

desired_order <- c("ICAR", "BW", "BW_2","RW_10", "EW_10","RW_20", "EW_20", 
                       "RW_50","EW_50", "RW_86", "EW_86", "RW_C", "EW_C")

# for WAIC

res_WAIC <- res_df[index_WAIC]

#removes the last "_xyz" from each colname, i.e. _DIC or _CPO
colnames(res_WAIC) <- sub("_(?!.*_).*", "", colnames(res_WAIC), perl = TRUE)

WAIC_long <-  res_WAIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "WAIC")

WAIC_long$Model <- factor(WAIC_long$Model, levels = desired_order)

WAIC_boxplot <- ggplot(WAIC_long, aes(x = Model, y = WAIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "WAIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# for DIC

res_DIC <- res_df[index_DIC]
colnames(res_DIC) <- sub("_(?!.*_).*", "", colnames(res_DIC), perl = TRUE) #chatGPT

DIC_long <-  res_DIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "DIC")

DIC_long$Model <- factor(DIC_long$Model, levels = desired_order)

DIC_boxplot <- ggplot(DIC_long, aes(x = Model, y = DIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "DIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# for CPO

res_CPO <- res_df[index_CPO]

colnames(res_CPO) <- sub("_(?!.*_).*", "", colnames(res_CPO), perl = TRUE) #chatGPT

CPO_long <-  res_CPO |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "CPO")

CPO_long$Model <- factor(CPO_long$Model, levels = desired_order)

CPO_boxplot <- ggplot(CPO_long, aes(x = Model, y = CPO)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "LS") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

full_boxplot <- ggarrange(WAIC_boxplot, DIC_boxplot, CPO_boxplot, nrow = 3)
full_boxplot
```


Plot for all diseases with ICAR as ref and with outliers.
```{r}
# for WAIC

res_WAIC <- res_df[index_WAIC]

colnames(res_WAIC) <- sub("_(?!.*_).*", "", colnames(res_WAIC), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_WAIC <- res_WAIC |>
  mutate(across(everything(), ~ .x - .data$ICAR))

WAIC_long <-  res_WAIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "WAIC")

WAIC_long$Model <- factor(WAIC_long$Model, levels = desired_order)

WAIC_boxplot <- ggplot(WAIC_long, aes(x = Model, y = WAIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "WAIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# for DIC

res_DIC <- res_df[index_DIC]
colnames(res_DIC) <- sub("_(?!.*_).*", "", colnames(res_DIC), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_DIC <- res_DIC |>
  mutate(across(everything(), ~ .x - .data$ICAR))

DIC_long <-  res_DIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "DIC")

DIC_long$Model <- factor(DIC_long$Model, levels = desired_order)

DIC_boxplot <- ggplot(DIC_long, aes(x = Model, y = DIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "DIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# for CPO

res_CPO <- res_df[index_CPO]

colnames(res_CPO) <- sub("_(?!.*_).*", "", colnames(res_CPO), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_CPO <- res_CPO |>
  mutate(across(everything(), ~ .x - .data$ICAR))

CPO_long <-  res_CPO |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "CPO")

CPO_long$Model <- factor(CPO_long$Model, levels = desired_order)

CPO_boxplot <- ggplot(CPO_long, aes(x = Model, y = CPO)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "LS") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

full_boxplot_W_outliers <- ggarrange(WAIC_boxplot, DIC_boxplot, CPO_boxplot, nrow = 3)
full_boxplot_W_outliers
```


Plot for all diseases with ICAR as reference and without showing outliers.
```{r}

# for WAIC

lower <- quantile(WAIC_long$WAIC, 0.025, na.rm = TRUE)
upper <- quantile(WAIC_long$WAIC, 0.975, na.rm = TRUE)

WAIC_boxplot <- ggplot(WAIC_long, aes(x = Model, y = WAIC)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "WAIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

# for DIC

lower <- quantile(DIC_long$DIC, 0.025, na.rm = TRUE)
upper <- quantile(DIC_long$DIC, 0.975, na.rm = TRUE)

DIC_boxplot <- ggplot(DIC_long, aes(x = Model, y = DIC)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "DIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

# for CPO

lower <- quantile(CPO_long$CPO, 0.025, na.rm = TRUE)
upper <- quantile(CPO_long$CPO, 0.975, na.rm = TRUE)

CPO_boxplot <- ggplot(CPO_long, aes(x = Model, y = CPO)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "LS") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

full_boxplot_WO_outliers <- ggarrange(WAIC_boxplot, DIC_boxplot, CPO_boxplot, nrow = 3)
full_boxplot_WO_outliers
```

Plot with only cancers, still ICAR as ref and no outliers.
```{r}

res_df <- res_df[7:33,] #only the cancer results

# for WAIC

res_WAIC <- res_df[index_WAIC]

colnames(res_WAIC) <- sub("_(?!.*_).*", "", colnames(res_WAIC), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_WAIC <- res_WAIC |>
  mutate(across(everything(), ~ .x - .data$ICAR))

WAIC_long <-  res_WAIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "WAIC")

WAIC_long$Model <- factor(WAIC_long$Model, levels = desired_order)

lower <- quantile(WAIC_long$WAIC, 0.025, na.rm = TRUE)
upper <- quantile(WAIC_long$WAIC, 0.975, na.rm = TRUE)

WAIC_boxplot <- ggplot(WAIC_long, aes(x = Model, y = WAIC)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "WAIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

# for DIC

res_DIC <- res_df[index_DIC]
colnames(res_DIC) <- sub("_(?!.*_).*", "", colnames(res_DIC), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_DIC <- res_DIC |>
  mutate(across(everything(), ~ .x - .data$ICAR))

DIC_long <-  res_DIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "DIC")

DIC_long$Model <- factor(DIC_long$Model, levels = desired_order)

lower <- quantile(DIC_long$DIC, 0.025, na.rm = TRUE)
upper <- quantile(DIC_long$DIC, 0.975, na.rm = TRUE)

DIC_boxplot <- ggplot(DIC_long, aes(x = Model, y = DIC)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "DIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

# for CPO

res_CPO <- res_df[index_CPO]

colnames(res_CPO) <- sub("_(?!.*_).*", "", colnames(res_CPO), perl = TRUE) #chatGPT

#Make the results with the standard ICAR as a reference level
res_CPO <- res_CPO |>
  mutate(across(everything(), ~ .x - .data$ICAR))

CPO_long <-  res_CPO |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "CPO")

CPO_long$Model <- factor(CPO_long$Model, levels = desired_order)

lower <- quantile(CPO_long$CPO, 0.025, na.rm = TRUE)
upper <- quantile(CPO_long$CPO, 0.975, na.rm = TRUE)

CPO_boxplot <- ggplot(CPO_long, aes(x = Model, y = CPO)) +
  geom_boxplot(outlier.shape = NA, fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "LS") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(lower, upper))

full_boxplot_cancer_WO_outliers <- ggarrange(WAIC_boxplot, DIC_boxplot, CPO_boxplot, nrow = 3)
full_boxplot_cancer_WO_outliers
```


```{r}
#Gamma prior
#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_W_outliers.png", plot = full_boxplot, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_ICAR_ref_W_outliers.png", plot = full_boxplot_W_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_ICAR_ref_WO_outliers.png", plot = full_boxplot_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_ICAR_ref_cancer_WO_outliers.png", plot = full_boxplot_cancer_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#PC prior
#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_PC_W_outliers.png", plot = full_boxplot, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_PC_ICAR_ref_W_outliers.png", plot = full_boxplot_W_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_PC_ICAR_ref_WO_outliers.png", plot = full_boxplot_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_PC_ICAR_ref_cancer_WO_outliers.png", plot = full_boxplot_cancer_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#Uniform prior
#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_U_W_outliers.png", plot = full_boxplot, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_U_ICAR_ref_W_outliers.png", plot = full_boxplot_W_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_U_ICAR_ref_WO_outliers.png", plot = full_boxplot_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)

#ggsave("Plots//Plots_withScaling_and_without_iidTheta_withCancer//CV_boxplot_U_ICAR_ref_cancer_WO_outliers.png", plot = full_boxplot_cancer_WO_outliers, height = 24, width = 21,  units = "cm",  dpi = 300)
```



