---
title: "Spatial Wakefield and geojson handeling"
author: "Halvard"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and processing the geosjon file
A geojson file for Spain and we are interested in the levels for autonomous regions and provinces. We descount the two islands, as they are less interesting in a spatial analysis. The goal is to make informative maps of the regions, as well as presenting the risk syrfaces and resutls of the analysis as a heatmap for the regions. For the analysis, we need the adjacency structure as a simple graph, or an adjacency matrix, and columns containing the name of the province and autonomous region in question. This must then also be harmonized with the disease data.

Some of the autonomous regions, like the island regions Illes Baleares and Canary Islands, and two cities in Morocco, namely Ceuta and Melilla, should be removed from all the data.

```{r}
library(sf)
library(spdep)
library(dplyr)
library(ggplot2)
library(ggpubr)

shp_prov <- st_read("Data/Provinces/Provincias_ETRS89_30N.shp")
colnames(shp_prov) <- c("Index", "ProvName", "ProvName1", "CCAA_Index", "CCAA_Name", "geometry")

#preporcessing of the data
CCAA_to_remove <- c("Illes Balears", "Canarias", "Ceuta", "Melilla")
shp_prov <- filter(shp_prov, !CCAA_Name %in% CCAA_to_remove)

plot(shp_prov["ProvName"], key.pos = NULL, main = "Map of the provinces of Spain")

shp_prov_large <- shp_prov |>
  group_by(CCAA_Name) |>
  summarise(geometry = st_union(geometry))

# Create the ggplot2 map
MapSpainFig <- ggplot() +
geom_sf(data = shp_prov, fill = NA, color = "black", linewidth = 0.6) +
geom_sf(data = shp_prov_large, fill = NA, color = "black", linewidth = 1.15) +
labs(title = "Map of autonomous regions and provinces in Spain") +
theme_minimal()
MapSpainFig

```


## Code for data preperation and plotting

```{r}
library(sf)
library(ggplot2)
library(dplyr)

edges_df <- readRDS("Edges_df//Edges_W_scaling_WO_theta")

ss_provinces = read_sf("Data/Provinces/Provincias_ETRS89_30N.shp")

centroids <- st_centroid(ss_provinces)

ss_ar = read_sf("Data/AutonomousRegions/Comunidades_Autonomas_ETRS89_30N.shp")

# we only work with mainland Spain, remove the islands

out = c("Islas Baleares", "Las Palmas", "Santa Cruz de Tenerife", "Ceuta", "Melilla")

out2 = c("Canarias", "Islas Baleares", "Ceuta", "Melilla")

ss_provinces = ss_provinces |> filter(!(Texto %in% out))

ss_ar = ss_ar |> filter(!(Texto %in% out2))

SpainMap <- ggplot() +
  geom_sf(data=ss_provinces, fill=NA, color="grey40") +
  geom_sf(data = ss_ar, color="black", fill=NA, linewidth = 1) +
  theme_minimal()

#ggsave("Plots//SpainMap.png", plot = SpainMap, height = 21, width = 21,  units = "cm",  dpi = 300)

# Map of Spain with illustrative graph for powerpoint
# SpainMap_W_graph <- ggplot() +
#   geom_sf(data = ss_provinces, fill = NA, color = "grey40") +  # Provinces
#   geom_sf(data = ss_ar, color="black", fill=NA,linewidth = 1) + # Autonomous regions
#   geom_sf(data = centroids, color = "red", size = 3) +  # Centroids
#   geom_sf(data = edges_df, linewidth = 0.8, color = "blue") + 
#   theme_minimal()

#ggsave("Plots//SpainMap_W_grpah.png", plot = SpainMap_W_graph, height = 21, width = 21,  units = "cm",  dpi = 300)

SpainMap_Wakefieldgraph <- ggplot() +
  geom_sf(data = ss_provinces, fill = NA, color = "grey40") +  # Provinces
  geom_sf(data = ss_ar, color="black", fill=NA,linewidth = 1) + # Autonomous regions
  geom_sf(data = centroids, color = "black", size = 3) +  # Centroids
  geom_sf(data = edges_df, linewidth = 1, aes(color = factor(inter_region_edges))) + 
  #scale_linewidth(range = c(0.5, 1.5)) +  
  scale_color_manual(values = c("1" = "blue", "0" = "red")) + 
  theme_minimal() +
  guides(linewidth = "none", color="none")

#ggsave("Plots//SpainMap_Wakefieldgraph.png", plot = SpainMap_Wakefieldgraph, height = 21, width = 21,  units = "cm",  dpi = 300)
```


# Making map figures for the report
First up is maps with edges to illustrate the Wakefield neighborhood structure. One map with grouping based on inside or across autonomous regions and one based on a "shocked" subset of provinces. Show one edge groups with red edges and the other as blue.

```{r}
# Figures to illustarte two choices for grouping the edges in the Border weighted ICAR

include <- c("15", "24", "27", "32", "33", "36", "49") 

include2 <- c("03", "12")

ss_provinces_example = filter(ss_provinces, Codigo %in% include)

ss_ar_example = filter(ss_ar, Codigo %in% include2)

centroids_example <- st_centroid(ss_provinces_example)

nb_example <- poly2nb(ss_provinces_example)

edge_list_example <- list()
region_from <- c()
region_to <- c()

for (i in seq_along(nb_example)) {
  neighbors <- nb_example[[i]]
  if (length(neighbors) == 0) next
  
  for (j in neighbors) {
    # Store the line geometry
    edge_list_example <- append(edge_list_example, 
                        list(st_linestring(rbind(st_coordinates(centroids_example[i, ]), 
                                                 st_coordinates(centroids_example[j, ])))))
    
    # Store region names or IDs
    region_from <- c(region_from, i)
    region_to <- c(region_to, j)
  }
}

edges_sf_example <- st_sf(geometry = st_sfc(edge_list_example, 
                                            crs = st_crs(ss_provinces_example)), 
                  from = region_from, to = region_to)

weights_example1 <- c(0, 0, 1, 1, 1, 0, 0, 1, 0, 1, 0, 1, 0, 0, 1, 1, 1, 0, 0, 0, 0, 1)

weights_example2 <- c(1, 0, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0, 1, 0, 0, 1, 1, 0, 1, 0, 0, 0)

edges_sf_example$Wex1 <- factor(weights_example1)
edges_sf_example$Wex2 <- factor(weights_example2)

ExamplePlot1 <- ggplot() +
  geom_sf(data = ss_ar_example, color="black") + #do not really need this line
  geom_sf(data=ss_provinces_example, color="grey20", linewidth = 0.5) +
  geom_sf(data = ss_ar_example, color="black", fill=NA,linewidth = 0.8) +
  geom_sf(data = centroids_example, color = "black", size = 2)  + # Centroids
  geom_sf(data = edges_sf_example, aes(color = Wex1), show.legend = FALSE) +
  scale_color_manual(values = c("0" = "blue", "1" = "red")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  theme_minimal()

#to highlight the two abnormal regions
ss_provinces_example_highlight_27 = filter(ss_provinces, Codigo %in% c(27))
ss_provinces_example_highlight_33 = filter(ss_provinces, Codigo %in% c(33))

ExamplePlot2 <- ggplot() +
  geom_sf(data = ss_ar_example, color="black") + #do not really need this line
  geom_sf(data=ss_provinces_example, color="grey20", linewidth = 0.5) +
  geom_sf(data = ss_ar_example, color="black", fill=NA,linewidth = 0.8) +
  geom_sf(data = ss_provinces_example_highlight_27, color = "black", 
          linewidth = 0.5, fill = "lightcoral") +
  geom_sf(data = ss_provinces_example_highlight_33, color = "black", 
          linewidth = 0.5, fill = "lightcoral") +
  geom_sf(data = centroids_example, color = "black", size = 2)  + # Centroids
  geom_sf(data = edges_sf_example, aes(color = Wex2), show.legend = FALSE) +
  scale_color_manual(values = c("0" = "blue", "1" = "red")) +
  theme(axis.text = element_blank(), axis.ticks = element_blank()) +
  theme_minimal()

examplePlots <- ggarrange(ExamplePlot1, ExamplePlot2, ncol = 2)

#ggsave("Plots//ExampleEdgesFigure.png", plot = examplePlots, height = 10, width = 21,  units = "cm",  dpi = 300)

#saving the individual plots

#ggsave("Plots//ExampleEdgesFigureA.png", plot = ExamplePlot1, height = 10, width = 10,  units = "cm",  dpi = 300)
#ggsave("Plots//ExampleEdgesFigureB.png", plot = ExamplePlot2, height = 10, width = 10,  units = "cm",  dpi = 300)

```

Making figures for the top-left autonomous region and the graph for the four provinces within it. 

```{r}
#Figure to show the corresponding graph with numbered nodes for the top left autonomous region of Spain

include <- c("15", "27", "32", "36")  #only prov in op left autonomous region of spain

include2 <- c("12")

ss_provinces_example = filter(ss_provinces, Codigo %in% include)

ss_ar_example = filter(ss_ar, Codigo %in% include2)

centroids_example <- st_centroid(ss_provinces_example)

nb_example <- poly2nb(ss_provinces_example)

edge_list_example <- list()
region_from <- c()
region_to <- c()

for (i in seq_along(nb_example)) {
  neighbors <- nb_example[[i]]
  if (length(neighbors) == 0) next
  
  for (j in neighbors) {
    # Store the line geometry
    edge_list_example <- append(edge_list_example, 
                        list(st_linestring(rbind(st_coordinates(centroids_example[i, ]), 
                                                 st_coordinates(centroids_example[j, ])))))
    
    # Store region names or IDs
    region_from <- c(region_from, i)
    region_to <- c(region_to, j)
  }
}

edges_sf_example <- st_sf(geometry = st_sfc(edge_list_example, 
                                            crs = st_crs(ss_provinces_example)), 
                  from = region_from, to = region_to)

weights_example1 <- rep(1, 10)


edges_sf_example$Wex21 <- factor(weights_example1)

centroids_example$region_ID <- c(1,2, 4, 3)

centroids_example_coords <- cbind(
  centroids_example,
  st_coordinates(centroids_example)
)

ExamplePlot1 <- ggplot() +
  #geom_sf(data = ss_ar_example, color="black") + #do not really need this line
  geom_sf(data = ss_provinces_example, fill = "grey80", color = "grey10", linewidth = 0.7) +
  geom_sf(data = ss_ar_example, color="black", fill=NA,linewidth = 0.8) +
  geom_sf(data = centroids_example, color = "red", size = 4)  + # Centroids
  geom_sf(data = edges_sf_example, aes(color = Wex21), linewidth = 1, show.legend = FALSE) +
  scale_color_manual(values = c("0" = "blue", "1" = "red")) +
   # Label using coordinates
  geom_text(data = centroids_example_coords,
            aes(x = X, y = Y, label = region_ID),  # Replace 'region_id' with your label column
            hjust = 1.5,  # vertical offset
            size = 8,
            color = "black") +
  coord_sf() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  theme(panel.background = element_rect(fill = "white", color = NA))

#ggsave("Plots//MapForGraphExample.png", plot = ExamplePlot1, height = 13, width = 13,  units = "cm",  dpi = 300)


```



# Spatial Wakefield model - Border weighted ICAR

The implementation will be in rgeneric and be very similar to the temporal case from my master project in the fall. The main difference is that the previous conflict and non-conflict points are now instead transitions inside an autonomous regions and transitions involving two autonomous regions. The expectation is that regions in the same autonomous regions have more in common than regions in different autonomous regions. Additionally, the function will need the dataframe with the administrative levels as an input variable to construct the precision matrix $Q$. Lets implement this with the dataframe created above as an input.

NB: There must be consistent ordering of regions and so on for the weight matrix to coincide with the supplied data. 

```{r}
library(INLA)

inla.rgeneric.BWICAR = function(
  cmd = c("graph", "Q", "mu", "initial", "log.norm.const","log.prior", "quit"),
  theta = NULL)
{
#Input:
#df: contains the relevant information for the areal data
#prior_str is either Gamma, PC1, PC3 or U
#AMat is the adjacency matrix, so 1 for bordering regions representing a first order structure
  
  envir = parent.env(environment())
  
  interpret_theta <- function() { return(list(tau1 = exp(theta[1L]), 
                                              tau2 = exp(theta[2L])))}
  
  graph <- function() {return(Q())}
  
  Q <- function() {
    N <- nrow(df) #df is passed as an argument, N is the number of regions
    R1 <- matrix(0, nrow = N, ncol = N) #non-conflict
    R2 <- matrix(0, nrow = N, ncol = N) #conflict
    
    #assume the adjacency matrix is called AMat, then
    non_zero_indices <- which(AMat == 1, arr.ind = TRUE)
    
    for (k in seq_len(nrow(non_zero_indices))) { #nrow?
      i <- non_zero_indices[k, 1]
      j <- non_zero_indices[k, 2]
      
      if(df[["Cod_CCAA"]][i] == df[["Cod_CCAA"]][j]){
        R1[i, j] <- -1
        R1[i, i] <- R1[i, i] + 1
      }
      else{
        R2[i, j] <- -1
        R2[i, i] <- R2[i, i] + 1
      }
    }
    
    gv <- exp(1 / N * sum(log(diag(INLA:::inla.ginv(R1 + R2))))) #scaling constant
    R_star_list <- list(R1 = R1*gv, R2 = R2*gv)
    
    p <- interpret_theta()
    Q <- R_star_list$R1 * p$tau1 + R_star_list$R2 * p$tau2
    return(inla.as.sparse(Q)) #sparse representation
  }
    
  mu <- function() {return(numeric(0))}
  
  initial <- function() {return(c(4, 4))}#Default initial for precisions is 4
  
  log.norm.const <- function() {return(numeric(0))}
  
  log.prior <- function() {#default: shape = 1, rate = 0.00005 for tau
    p <- interpret_theta()
    if (pr_str == "gamma")  {
      prior <- dgamma(p$tau1, shape = 1, rate = 0.00005, log = TRUE) + log(p$tau1) +
        dgamma(p$tau2, shape = 1, rate = 0.00005, log = TRUE) + log(p$tau2)}
    else if (pr_str == "PC3") {
      prior <- inla.pc.dprec(p$tau1, u = 3, alpha = 0.05, log=TRUE) + log(p$tau1) +
        inla.pc.dprec(p$tau2, u = 3, alpha = 0.05, log=TRUE) + log(p$tau2)}
    else if (pr_str == "PC1") {
      prior <- inla.pc.dprec(p$tau1, u = 1, alpha = 0.01, log=TRUE) + log(p$tau1) +
        inla.pc.dprec(p$tau2, u = 1, alpha = 0.01, log=TRUE) + log(p$tau2)}
    else if (pr_str == "U") {prior <- -0.5*log(p$tau1) - 0.5*log(p$tau2)}
    return(prior)
  }
  
  quit <- function() {return(invisible())}
  
  #to ensure theta is defined
  if (!length(theta)) theta = initial()
  
  vals <- do.call(match.arg(cmd), args = list())
  return(vals)
}
```

The following is used to initialize the Border weighted model inside the loop in the chuck below.
```{r}
nb <- poly2nb(ss_provinces)
adj_matrix <- nb2mat(nb, style = "B", zero.policy = TRUE)
nAreas <- nrow(adj_matrix)
```

## Sensitivity analysis for BW-ICAR
Comparison of four different priors for the two precisions in BW. 

```{r, eval=FALSE}
#Load the data
ExpectedCases <- readRDS("ExpectedCasesMiguel.rds")
ObservedCases <- readRDS("ObservedCases.rds")

#Which diseases to include
test_indecies <- 1:86

sensitivity_df <- data.frame(matrix(NA, nrow = length(test_indecies), ncol = 12))
colnames(sensitivity_df) <- c("gamma_WAIC", "PC1_WAIC", "PC3_WAIC", "U_WAIC", 
                              "gamma_DIC", "PC1_DIC", "PC3_DIC", "U_DIC",
                              "gamma_CPO", "PC1_CPO", "PC3_CPO", "U_CPO")

prior_options <- c("gamma", "PC1", "PC3", "U")

for (i in 1:4){
  BW_sens_model <- inla.rgeneric.define(inla.rgeneric.BWICAR, df = ss_provinces, 
                                    AMat = adj_matrix, pr_str = prior_options[i]) 
  for (m in 1:86){
    Y <- as.vector(ObservedCases[, j])
    E <- as.vector(ExpectedCases[, j])
  
    data_df <- data.frame(Y = Y, E = E, ID = 1:nAreas)
    
    formula_BW <- Y ~ offset(log(E)) + f(ID, model = BW_sens_model)
    res_BW <- inla(formula_BW, family = "poisson", data = data_df,
                 control.compute = list(cpo = T, waic = T, dic = T))
    
    sensitivity_df[m, i] <- res_BW$waic$waic
    sensitivity_df[m, i + 4] <- res_BW$dic$dic
    sensitivity_df[m, i + 8] <- -mean(log(res_BW$cpo$cpo))
  }
}

#save the results
#saveRDS(sensitivity_df, file = "Results//BW-sensitivity.rds")
```

Visualizing the results:
```{r}
library(tidyr)
library(dplyr)

sensitivity_df <- readRDS(file = "Results//BW-sensitivity.rds")

index_WAIC <- 1:4

index_DIC <- 5:8

index_CPO <- 9:12

#For WAIC
sens_WAIC <- sensitivity_df[index_WAIC]

colnames(sens_WAIC) <- sub("_(?!.*_).*", "", colnames(sens_WAIC), perl = TRUE)

#transform to PC3 as a referance level
sens_WAIC <- sens_WAIC |>
  mutate(across(everything(), ~ .x - .data$PC3))

WAIC_long <-  sens_WAIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "WAIC")

WAIC_boxplot <- ggplot(WAIC_long, aes(x = Model, y = WAIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "WAIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#For DIC
sens_DIC <- sensitivity_df[index_DIC]

colnames(sens_DIC) <- sub("_(?!.*_).*", "", colnames(sens_DIC), perl = TRUE)

#transform to PC3 as a referance level
sens_DIC <- sens_DIC |>
  mutate(across(everything(), ~ .x - .data$PC3))

DIC_long <-  sens_DIC |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "DIC")

DIC_boxplot <- ggplot(DIC_long, aes(x = Model, y = DIC)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "DIC") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#For CPO
sens_CPO <- sensitivity_df[index_CPO]

colnames(sens_CPO) <- sub("_(?!.*_).*", "", colnames(sens_CPO), perl = TRUE)

#transform to PC3 as a referance level
sens_CPO <- sens_CPO |>
  mutate(across(everything(), ~ .x - .data$PC3))

CPO_long <-  sens_CPO |>
  mutate(row_id = row_number()) |>
  pivot_longer(-row_id, names_to = "Model", values_to = "CPO")

CPO_boxplot <- ggplot(CPO_long, aes(x = Model, y = CPO)) +
  geom_boxplot(fill = "skyblue") +
  theme_minimal() +
  labs(x = "", y = "CPO") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#saving the figures
# ggsave("Plots//SenitivityFigureWAIC.png", plot = WAIC_boxplot, height = 7, width = 10,  units = "cm",  dpi = 300)
# ggsave("Plots//SenitivityFigureDIC.png", plot = DIC_boxplot, height = 7, width = 10,  units = "cm",  dpi = 300)
# ggsave("Plots//SenitivityFigureCPO.png", plot = CPO_boxplot, height = 7, width = 10,  units = "cm",  dpi = 300)

```


# Plot for different priors on a precision tau
Specifically a gamma, PC and uniform on sd sigma prior.

```{r}

#gamma

x_gamma <- seq(1/25, 100000, length.out = 10000)

y_gamma <- dgamma(x_gamma, shape = 1, rate = 0.00005)

df <- data.frame(x = x_gamma, y = y_gamma)

min_x <- 0
max_x <- 100000

plot_gamma <- ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue", size = 1.2) +
  labs(x = expression(tau), y = expression(pi(tau))) +
  scale_x_continuous(breaks = seq(min_x, max_x, length.out = 3)) +
  theme_minimal(base_size = 14)

#gamma2

x_gamma2 <- seq(1/25, 2, length.out = 10000)

y_gamma2 <- dgamma(x_gamma2, shape = 1, rate = 0.00005)

df <- data.frame(x = x_gamma2, y = y_gamma2)

plot_gamma2 <- ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue", size = 1.2) +
  labs(x = expression(tau), y = expression(pi(tau))) +
  theme_minimal(base_size = 14)


#uniform on sigma

x_uniform <- seq(1/25, 1, length.out = 10000)

U_function <- function(x){
  return(0.1*x**(-3/2))
}

y_uniform <- U_function(x_uniform)

df <- data.frame(x = x_uniform, y = y_uniform)

plot_U <- ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue", size = 1.2) +
  labs(x = expression(tau), y = expression(pi(tau))) +
  theme_minimal(base_size = 14)

#PC

x_PC <- seq(1/25, 2, length.out = 10000)

u <- 3
alpha <- 0.05
lambda <- -log(alpha) / u

y_PC <- (lambda / 2) * x_PC**(-3/2) * exp(-lambda / sqrt(x_PC))

df <- data.frame(x = x_PC, y = y_PC)

plot_PC <- ggplot(df, aes(x = x, y = y)) +
  geom_line(color = "steelblue", size = 1.2) +
  labs(x = expression(tau), y = expression(pi(tau))) +
  theme_minimal(base_size = 14)


#saving the plots

# ggsave("Plots//Prior Example gamma.png", plot = plot_gamma, height = 8, width = 8, units = "cm",  dpi = 300)
# ggsave("Plots//Prior Example gamma [0, 2].png", plot = plot_gamma2, height = 8, width = 8, units = "cm",  dpi = 300)
# ggsave("Plots//Prior Example uniform.png", plot = plot_U, height = 8, width = 8, units = "cm",  dpi = 300)
# ggsave("Plots//Prior Example PC.png", plot = plot_PC, height = 8, width = 8, units = "cm",  dpi = 300)

```



# Load and preprocess the disease data
The data for male mortality cases nationally and at province level is collected in 5-year age bins for each disease. These must be aggregated to the total cases per disease in each province and summarized in a matrix called ObservedCases. Furthermore, we calculate the expected cases based on the population per age bin in each province, which is denoted as ExpectedCases. Lastly, some choice of the mortality causes must be done. For instance, some causes are extremely rare, for instance only 1 or 2 cases nationally. Therefore, we implemented a lower limit of an average of 2 cases per province, which for 47 provinces equals 94 cases nationally. Additionally, some of the mortality causes are not disease related, for instance homicides, but they are kept in the data. 

```{r}
library(readr)
library(tidyr)

#Calculation of expected observation

# load("ExpectedCases.Rdata")
# 
# colnames(Expected) <- gsub("^\\d+\\s+", "", colnames(Expected))
# rownames(Expected) <- gsub("^\\d+\\s+", "", rownames(Expected))
# 
# Expected2 <- as.data.frame(Expected)
# 
# saveRDS(Expected2, "ExpectedCasesMiguel.rds")

#Mortality data

# Mortality_df <- read_csv2("Mortality.csv")
# 
# Mortality_df <- Mortality_df[, c(1, 3, 4)]
# colnames <- colnames(Mortality_df)
# 
# df_wide <- Mortality_df %>%
#   pivot_wider(names_from = colnames[1], values_from = colnames[3], values_fill = list(cases = NA))
# df_wide <- df_wide[, -43] #remove a joint column
# 
# df_matrix <- as.matrix(df_wide[, -1]) #remove the province column
#
# ObservedCases <- as.matrix(ExpectedCases) #I made sure everything aligned, both rows and cols
# 
# ObservedCases[1:47, 1:102] <- df_matrix[1:47, 1:102]
# 
# sparse_columns <- as.numeric(which(colSums(df_matrix) < 94)) # 47*2=94, was 16 columns
# 
# ObservedCases <- ObservedCases[, -sparse_columns] #removed the 16 sparse columns
# 
# saveRDS(ObservedCases, "ObservedCases.rds")
# 
# ObservedCases <- readRDS("ObservedCases.rds")
# 
#ExpectedCases <- readRDS("ExpectedCasesMiguel.rds")
#
#ExpectedCases <- as.matrix(ExpectedCases)
# 
# ExpectedCases <- ExpectedCases[, -sparse_columns]
# 
# saveRDS(ExpectedCases, "ExpectedCasesMiguel.rds")

#Population data
# Pop_data <- read_csv2("PopXProvincesXAge.csv")
# 
# colnames(Pop_data)
# 
# library(dplyr)
# 
# Pop_data_total <- Pop_data |>
#   group_by(Provinces) |>  # Group by Provinces
#   summarise(Total_Population = sum(Total, na.rm = TRUE))  # Sum Population column
# 
# saveRDS(Pop_data_total, "PopulationData.rds")
```



