---
title: "Temporal simulation"
author: "Halvard"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Temporal simulations
This file will create simulations for the four temporal models mentioned in the theory introudction of my thesis. Specificially, RW1, RW2, AR1 and IID.

## RW1
The RW1 follows the following dependence structure:
$$
x_{t+1} \sim N(x_t, \sigma^2), \quad for t = 1, ..., T-1 
$$

```{r}
RW1 <- function(sigma, N){
  # sigma^2 is the variance for the transitions and N is the number of points
  x <- rep(0, N)
  z <- sigma * rnorm(N-1)
  for(j in 2:N){
    x[j] <- x[j-1] + z[j-1]
  }
  return(x)
}
```

## RW2

```{r}
RW2 <- function(sigma, N){
  # sigma^2 is the variance for the transitions and N is the number of points
  x <- rep(0, N)
  z <- sigma * rnorm(N-2)
  for(j in 3:N){
    x[j] <- 2*x[j-1] - x[j-2] + z[j-2]
  }
  return(x)
}
```

## AR1

```{r}

AR1 <- function(sigma, N, rho){
  # sigma^2 is the variance for the transitions, N is the number of points and rho is the AR parameter
  x <- rep(0, N)
  z <- sigma * rnorm(N-1)
  for(j in 2:N){
    x[j] <- rho*x[j-1] + z[j-1]
  }
  return(x)
}

```

## IID

```{r}

IID <- function(sigma, N){
  # sigma^2 is the variance for the transitions and N is the number of points
  z <- sigma * rnorm(N)
  return(z)
}

```


## Simulating and combining the plots

```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(ggpubr)

N <- 50
sigma <- 1
rho <- 0.8

#RW1
df_RW1 <- data.frame(col1 = RW1(sigma, N), col2 = RW1(sigma, N), col3 = RW1(sigma, N), x = 1:N)

df_long_RW1 <- pivot_longer(df_RW1, cols = c(col1, col2, col3), names_to = "variable", values_to = "value")

plot_RW1 <- ggplot(df_long_RW1, aes(x = x, y = value, color = variable)) +
  geom_line() +
  labs(title = "RW1", x = "t", y = "y") +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

#RW2
df_RW2 <- data.frame(col1 = RW2(sigma, N), col2 = RW2(sigma, N), col3 = RW2(sigma, N), x = 1:N)

df_long_RW2 <- pivot_longer(df_RW2, cols = c(col1, col2, col3), names_to = "variable", values_to = "value")

plot_RW2 <- ggplot(df_long_RW2, aes(x = x, y = value, color = variable)) +
  geom_line() +
  labs(title = "RW2", x = "t", y = "y") +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

#AR1
df_AR1 <- data.frame(col1 = AR1(sigma, N, rho), col2 = AR1(sigma, N, rho), 
                     col3 = AR1(sigma, N, rho), x = 1:N)

df_long_AR1 <- pivot_longer(df_AR1, cols = c(col1, col2, col3), names_to = "variable", values_to = "value")

plot_AR1 <- ggplot(df_long_AR1, aes(x = x, y = value, color = variable)) +
  geom_line() +
  labs(title = "AR1", x = "t", y = "y") +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

#IID
df_IID <- data.frame(col1 = IID(sigma, N), col2 = IID(sigma, N), col3 = IID(sigma, N), x = 1:N)

df_long_IID <- pivot_longer(df_IID, cols = c(col1, col2, col3), names_to = "variable", values_to = "value")

plot_IID <- ggplot(df_long_IID, aes(x = x, y = value, color = variable)) +
  geom_line() +
  labs(title = "IID", x = "t", y = "y") +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

#Final figure

temporal_figure <- ggarrange(plot_RW1, plot_RW2, plot_AR1, plot_IID, nrow = 2, ncol = 2)

#ggsave("Plots//Temporal_figure.png", plot = temporal_figure, height = 21, width = 21,  units = "cm",  dpi = 300)
```


# Linear regression example with priors
A simple example to show prior importance in a simple linear regression. The model is
$$
y \sim N(\eta, 0.0000001^2) \\
\eta = \mu + \beta x
$$
with some priors on mu and beta. Lets say the true beta is 5 and mu is 100

```{r}
x <- c(10, 20, 30, 40, 50)
y <- c(135, 210, 240, 280, 370)

df <- data.frame(x = x, y = y)

formula <- "y ~ x"

lm_res <- lm(formula, df)

summary(lm_res)
```

INLA model:
```{r}
library(INLA)
INLA_formula <- y ~ 1 + x

res_standard <- inla(INLA_formula, data = df, control.fixed = list(
    mean.intercept = 0,      # Prior mean for intercept (μ)
    prec.intercept = 1e-6,   # Prior precision for intercept
    mean = 0,                # Prior mean for slope (β)
    prec = 1e-6              # Prior precision for slope
  ))

summary(res_standard)

df$standard <- c(res_standard$summary.linear.predictor$mean)

res_bad_mu <- inla(INLA_formula, data = df, control.fixed = list(
    mean.intercept = 0,      # Prior mean for intercept (μ)
    prec.intercept = 1e2,   # Prior precision for intercept
    mean = 0,                # Prior mean for slope (β)
    prec = 1e-6              # Prior precision for slope
  ))

summary(res_bad_mu)

df$bad_mu <- c(res_bad_mu$summary.linear.predictor$mean)

res_bad_beta <- inla(INLA_formula, data = df, control.fixed = list(
    mean.intercept = 0,      # Prior mean for intercept (μ)
    prec.intercept = 1e-6,   # Prior precision for intercept
    mean = 2,                # Prior mean for slope (β)
    prec = 1e2              # Prior precision for slope
  ))

summary(res_bad_beta)

df$bad_beta <- c(res_bad_beta$summary.linear.predictor$mean)

res_bad_both <- inla(INLA_formula, data = df, control.fixed = list(
    mean.intercept = 0,      # Prior mean for intercept (μ)
    prec.intercept = 1e2,   # Prior precision for intercept
    mean = 2,                # Prior mean for slope (β)
    prec = 1e2              # Prior precision for slope
  ))

summary(res_bad_both)

df$both_bad <- c(res_bad_both$summary.linear.predictor$mean)
```

Plotting:
```{r}
stand_plot <- ggplot(df, aes(x = x)) +
  geom_point(aes(y = y), color = "blue") +
  geom_line(aes(y = standard), color = "red", linewidth = 1) +
  labs(title = expression(mu %~% N(0, 10^6) * " and " * beta %~% N(0, 10^6)), 
       x = "x", y = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18),
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14))

bad_mu_plot <- ggplot(df, aes(x = x)) +
  geom_point(aes(y = y), color = "blue") +
  geom_line(aes(y = bad_mu), color = "red", linewidth = 1) +
  labs(title = expression(mu %~% N(0, 10^{-2}) * " and " * beta %~% N(0, 10^6)), 
       x = "x", y = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18),
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14))

bad_beta_plot <- ggplot(df, aes(x = x)) +
  geom_point(aes(y = y), color = "blue") +
  geom_line(aes(y = bad_beta), color = "red", linewidth = 1) +
  labs(title = expression(mu %~% N(0, 10^6) * " and " * beta %~% N(2, 10^{-2})), 
       x = "x", y = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18),
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14))

both_bad_plot <- ggplot(df, aes(x = x)) +
  geom_point(aes(y = y), color = "blue") +
  geom_line(aes(y = both_bad), color = "red", linewidth = 1) +
  labs(title = expression(mu %~% N(0, 10^{-2}) * " and " * beta %~% N(2, 10^{-2})), 
       x = "x", y = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18),
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14))

full_plot <- ggarrange(stand_plot, bad_mu_plot, bad_beta_plot, both_bad_plot, ncol = 2, nrow = 2)

#ggsave("Plots//linearRegressionWPriors.png", plot = full_plot, height = 21, width = 21,  units = "cm",  dpi = 300)

```

# Polynomial example with IID
The idea is to compare models with and without iid, and compare different degrees of polynomials.

```{r}
x <- c(1, 2, 3, 4, 5, 6, 7, 8)#, 9, 10)
y <- c(6, 7, 13, 15, 25, 23, 29, 35)#, 36, 44)

df2 <- data.frame(x = x, y = y)


res_1poly <- lm(y ~ poly(x, 1, raw = TRUE), data = df2)

res_5poly <- lm(y ~ poly(x, 7, raw = TRUE), data = df2)


x_grid <- data.frame(x = seq(min(df2$x), max(df2$x), length.out = 200))
x_grid$y_1 <- predict(res_1poly, newdata = x_grid)
x_grid$y_5 <- predict(res_5poly, newdata = x_grid)

linRegPlot <- ggplot(df2, aes(x, y)) +
  geom_point(color = "blue", size = 2.5) +
  geom_line(data = x_grid, aes(x = x, y = y_1), color = "red", size = 1, linetype = "longdash") +
  geom_line(data = x_grid, aes(x = x, y = y_5), color = "red", size = 1) +
  labs(x = "t", y = "y") +
  theme_minimal()

#ggsave("Plots//PolyFig.png", plot = linRegPlot, height = 7, width = 10,  units = "cm",  dpi = 300)

```




